Solutions to Integer Programming from Quantum Annealing

### 2020-05-26 paper/main.tex

%----------------------------------------------------------------------------------------
\subsubsection{Minimum Dominating Set Problem}
\label{sec:methods:mds-def}
%----------------------------------------------------------------------------------------

@@

The target Hamiltonian can be written in the matrix form
\begin{widetext}
\begin{align}
    E = &
    \begin{pmatrix}
    \Psi^x & \Psi^s
    \end{pmatrix}
    \begin{pmatrix}
    \mathbbm{1} + p\left[J^T J + J^T + J - 2 \mathrm{Diag}_{\psi^x}(|J|) - \mathbbm{1} \right] & - p(\mathbbm{1}+J^T)T^s\\
    - p{T^s}^T(\mathbbm{1}+J)& p\left[{T^s}^T T^s + 2\mathrm{Diag}_{\psi^s}(|T^s|)\right]
    \end{pmatrix}
    \begin{pmatrix}
    \Psi^x\\ \Psi^s
    \end{pmatrix} + p \overline{\overline{V}}
    \label{eq:matrix_form}
\end{align}
\end{widetext}
where $ |M| \equiv \sum_{\nu} M_{\nu \mu}$.

@@

In quantum annealing, the ground state of the problem Hamiltonian is obtained from the evolution of a time-dependent Hamiltonian,
\begin{equation}
    H(s) = A(s) H^{\textrm{init}} + B(s) H^{\textrm{problem}}, \label{eq:tdhamiltonian}
\end{equation}
where $H^\textrm{init}=\sum_i\sigma^x_i$ on the DWave, and $H^\textrm{problem}$ is given by Eq.~(\ref{eq:matrix_form}) for the MDS application.
The variable $s\in [0, 1]$ is the normalized annealing time and is allowed to vary between 1$\mu s$ to 2$ms$ (CK: time unit mapping) on the DWave.
On DWave solvers, annealing offsets effectively advance or delay the annealing schedule of individual qubits.
In Fig.~\ref{fig:anneal_schedule}, the default DWave annealing schedule is shown in addition to the effects of applying positive and negative offsets.
Note that due to the way the offsets are implemented on the hardware level, systematic errors are introduced since the initial and final values of $A(s)$ and $B(s)$ will differ for qubits with different offsets.
Fig.~\ref{fig:anneal_schedule} shows the maximum offset that we use for the study, which translates to approximately (due to non-linearity) a 5\% advancement (blue) or delay (red) in the schedule.

\begin{figure}[b]
    \centering
    \includegraphics[width=\columnwidth]{./figures/anneal_schedule.pdf}
    \caption{asdf}
    \label{fig:anneal_schedule}
\end{figure}

The motivation behind implementing per-qubit level offsets is rooted in the hypothesis that as the problem Hamiltonian is turned on, the system in general will exhibit more disorder.
We define disorder by first recasting the QUBO into its Ising form such that the problem Hamiltonian has the form
\begin{equation}
    H^{\textrm{Ising}} = \sum_{ij} J_{ij} \sigma^z_i \sigma^z_j + \sum_i h_i \sigma^z_i,
\end{equation}
and recognize that in the limit where $h_i$ is randomly drawn from a Bernoulli distribution we recover the spin-glass model.
The spin-glass model enters its glassy state when $|h_i|$ is large, and as a consequence, the wavefunction experiences many-body localization effects, the many-body analogue of Anderson Localization.

@@

%----------------------------------------------------------------------------------------
\subsubsection{The Lindblad Equation}
\label{sec:methods:lindblad}
%----------------------------------------------------------------------------------------
Say what the simulator is solving

{\color{blue}

To solve for time evolution dynamics of quantum annealing including thermal and the decoherence effects, we solve for the master equation in Lindblad form
\begin{align}
    \partial_t \rho (t) =  \frac{-i}{\hbar} [H(t) , \rho(t)] + \sum_j (2L_j \rho(t) L_j^\dagger - \{ L^\dagger_j L_j, \rho(t) \}) ,
\end{align}
where $\rho (t)$ is the density matrix at time $t$.
$H(t)$ is the time-dependent Hamiltonian
\begin{align}
    \label{eq:annealH}
     H_{anneal}(t)&  =  - \sum_i  A_i(t)\sigma^x_i & \notag \\
     +&  \sum_{i>j} \sqrt{B_{i}(t)B_{j}(t)} J_{ij} \sigma^z_i \sigma^z_j +\sum_i B_{i}(t) h_i \sigma^z_i  ,&
\end{align}
where $A_i(t)$ and $B_{i}(t)$ are site-dependent annealing schedule functions.
The site dependency takes into account of the annealing offset.
$[,]$ denotes the Lie bracket.
$L_j$ are the Lindblad operators for decoherence.
$\{, \}$ denotes the anti-commutator.
In this study we only include a minimal model for decoherence, i.e., the amplitude damping for non-interacting qubits.
The Lindblad operator is $L_j=\sqrt{\gamma} \sigma^{+}_j$ for $h_j>0$ and $L_j=\sqrt{\gamma} \sigma^{-}_j$ for $h_j<0$, where $\gamma = 1/T_1$ is the decoherence rate for relaxation time $T_1$.
For simplicity we assume all qubits having the same relaxation time.
The initial condition is the Gibbs canonical ensemble
\begin{equation}
    \rho (0) =  \frac{e^{-\beta H(0)}}{\mbox{Tr}[e^{-\beta H(0)}]} ,
\end{equation}
where $\beta$ is the inverse temperature.
The probability to get the ground state at measurement is
\begin{align}
    P =  \mbox{Tr} \{  \rho (t) \pi_{\mbox{gnd}} \}  ,
\end{align}
where the projection operator onto the degenerated ground states subspace is defined as $\pi_{\mbox{gnd}}=\sum_{i\in G} |\mbox{gnd}_i\rangle \langle \mbox{gnd}_i| $.
Here $\{ | \mbox{gnd}_i \rangle | i \in G \}$ forms an orthonormal basis for the degenerated ground states subspace, i.e., $\langle \mbox{gnd}_j | \mbox{gnd}_i \rangle = \delta_{ij}$.

}

With anneal offsets
%----------------------------------------------------------------------------------------
\subsubsection{Other Simulation Details}
%----------------------------------------------------------------------------------------
Temperature
Initial wave function choices
RK tolerance

>>>>

This sections demonstrates the formalism in case of an example ILP: the Minimum Dominating Set problem.

%----------------------------------------------------------------------------------------
\subsubsection{Definition of the Minimum Dominating Set Problem}
\label{sec:methods:mds-def}
%----------------------------------------------------------------------------------------

@@

The target QUBO in the notation of eq.~\eqref{eq:matrix_form} reads
{\small
\begin{align}
    Q_{xx} &= \mathbbm{1} + p\left[J^T J + J^T + J - 2 \mathrm{Diag}_{\psi^x}(|J|) - \mathbbm{1} \right] \,,\\
    Q_{sx} &= - p(\mathbbm{1}+J^T)T^s\,,\\
    Q_{ss} &= p\left[{T^s}^T T^s + 2\mathrm{Diag}_{\psi^s}(|T^s|)\right]\,,\\
    C &=  p \overline{\overline{V}}\,,
\end{align}}
where $ |M| \equiv \sum_{\nu} M_{\nu \mu}$.

@@

A quantum annealer evolves the ground state of an initial Hamiltonian into the ground state of the problem Hamiltonian by adiabatically changing the hardware which implements Hamiltonian.
Thus, the implemented Hamiltonian is explicitly time dependent
\begin{equation}
    H(s) = A(s) H^{\textrm{init}} + B(s) H^{\textrm{problem}}, \label{eq:tdhamiltonian}
\end{equation}
and $H^\textrm{init}=\sum_i\sigma^x_i$ on the DWave, while $H^\textrm{problem}$ is given by Eq.~(\ref{eq:matrix_form}) for the MDS application.
The variable $s\in [0, 1]$ is the normalized annealing time and, in physical units, is allowed to vary between 1$\mu s$ to 2$ms$ on DWave.

On DWave solvers, annealing offsets effectively advance or delay the annealing schedule of individual qubits ($B(s) \to B_i(s)$).
In Fig.~\ref{fig:anneal_schedule}, the default DWave annealing schedule is shown in addition to the effects of applying positive and negative offsets.
Note that due to the way the offsets are implemented on the hardware level, systematic errors are introduced since the initial and final values of $A(s)$ and $B(s)$ will differ for qubits with different offsets ($B_i(s) \neq B_j(s)$).
Fig.~\ref{fig:anneal_schedule} shows the maximum offset that we use for the study, which translates to approximately (due to non-linearity) a 5\% advancement (blue) or delay (red) in the schedule.

\begin{figure}[htb]
    \centering
    \includegraphics[width=\columnwidth]{./figures/anneal_schedule.pdf}
    \caption{asdf}
    \label{fig:anneal_schedule}
\end{figure}

The motivation behind implementing per-qubit level offsets is rooted in the hypothesis that as the problem Hamiltonian is turned on, the system in general will exhibit more disorder.
When recasting the QUBO into its Ising form such that the problem Hamiltonian has the form
\begin{equation}
    H^{\textrm{Ising}} = \sum_{ij} J_{ij} \sigma^z_i \sigma^z_j + \sum_i h_i \sigma^z_i,
\end{equation}
we recognize that if $h_i$ is randomly drawn from a Bernoulli distribution we recover the spin-glass model.
The spin-glass model enters its glassy state when $|h_i|$ is large, and as a consequence, the wavefunction experiences many-body localization effects, the many-body analogue of Anderson Localization.

@@

%----------------------------------------------------------------------------------------
\subsubsection{The Lindblad Equation}
\label{sec:methods:lindblad}
%----------------------------------------------------------------------------------------

To solve for time evolution dynamics of quantum annealing including thermal and the decoherence effects, we solve for the master equation in Lindblad form
\begin{align}
    \partial_t \rho (t) =  \frac{-i}{\hbar} [H(t) , \rho(t)] + \sum_j (2L_j \rho(t) L_j^\dagger - \{ L^\dagger_j L_j, \rho(t) \}) ,
\end{align}
where $\rho (t)$ is the density matrix at time $t$.
$H(t)$ is the time-dependent Hamiltonian
\begin{align}
    \label{eq:annealH}
     H_{anneal}(t)&  =  - \sum_i  A_i(t)\sigma^x_i & \notag \\
     +&  \sum_{i>j} \sqrt{B_{i}(t)B_{j}(t)} J_{ij} \sigma^z_i \sigma^z_j +\sum_i B_{i}(t) h_i \sigma^z_i  ,&
\end{align}
where $A_i(t)$ and $B_{i}(t)$ are site-dependent annealing schedule functions.
The site dependency takes into account of the annealing offset.
$[,]$ denotes the Lie bracket.
$L_j$ are the Lindblad operators for decoherence.
$\{, \}$ denotes the anti-commutator.
In this study, we only include a minimal model for decoherence, i.e., the amplitude damping for non-interacting qubits.
The Lindblad operator is $L_j=\sqrt{\gamma} \sigma^{+}_j$ for $h_j>0$ and $L_j=\sqrt{\gamma} \sigma^{-}_j$ for $h_j<0$, where $\gamma = 1/T_1$ is the decoherence rate for relaxation time $T_1$.
For simplicity, we assume all qubits have the same relaxation time.
The initial condition is the Gibbs canonical ensemble
\begin{equation}
    \rho (0) =  \frac{e^{-\beta H(0)}}{\mbox{Tr}[e^{-\beta H(0)}]} ,
\end{equation}
where $\beta$ is the inverse temperature.
The probability to obtain the ground state at for a single measurement is
\begin{align}
    P_{\mbox{gnd}} =  \mbox{Tr} \{  \rho (t) \pi_{\mbox{gnd}} \}  ,
\end{align}
where the projection operator onto the degenerated ground states subspace is defined as
\begin{equation}
	\pi_{\mbox{gnd}}=\sum_{i\in G} |\mbox{gnd}_i\rangle \langle \mbox{gnd}_i|
	\, .
\end{equation}
Here $\{ | \mbox{gnd}_i \rangle | i \in G \}$ forms an orthonormal basis for the degenerated ground states subspace, i.e., $\langle \mbox{gnd}_j | \mbox{gnd}_i \rangle = \delta_{ij}$.

%----------------------------------------------------------------------------------------
\subsubsection{Other Simulation Details}
%----------------------------------------------------------------------------------------
Temperature
Initial wave function choices
RK tolerance
